---
title: "Sandwich Sales Optimization"
author: "Partha Banerjee, Santiago Cetrangolo, Aaron Palumbo, Vincent Ying"
date: "Sunday, September 14, 2014"
output: html_document
---

James is a budding entrepreneur, and for the last two years he has been operating a sandwich stand 
in the lobby of his office building during the lunch hour. He also knows that careful data analysis 
can help him run his business more effectively. As a result, he has been tracking sandwich demand 
over the two years and he has carefully recorded the number of each type of sandwich demanded, 
the number he brought with him to sell, and his prices for each type of sandwich. 
Your job is to use the data to determine how many sandwiches of each type he should bring each 
day in order to maximize his expected profits. Two CSV files are available to support your analysis. 
The daily sales data contains number brought and number demanded for each sandwich. (Note that 
there may be days where demand exceeds supply, but James is clever and he records the requests 
he is unable to fulfill.) The pricing data contains the cost and sale price of each type of sandwich. 
Analyze the data and provide a recommendation to James. Be sure to identify any assumptions you 
make and be sure to back up your recommendations with thorough analysis. You should, as a 
group, submit a single report (expected to be 1-3 pages, but that is a guideline and not a rule) that 
provides your analysis. You may also attach any relevant code, but that is not required and will not 
be taken into account for grading purposes. 
 
 
This assignment should be submitted through the module 2 collaborative assignment mechanism in 
blackboard by end of day (11:59 p.m. EDT) on Wednesday, September 27, 2014. Each group has 
its own discussion board and file share available. 

```{r echo=FALSE}

# Make sure we're in the right working directory:  "Code"

if("Code" %in% list.files()){
  setwd("Code")
}

if(!("Code" %in% list.files("..")))
{
  stop
}
```

# Load Data

```{r data}
# Start by loading the data

sales <- read.csv("../Data/sales.csv")
details <- read.csv("../data/details.csv")
```

# Clean Data

```{r}
# Add variable for total demand
sales$demand.total <- sales$demand.ham + sales$demand.turkey + sales$demand.veggie

# Add variable for profit = price - cost
details$profit <- details$price - details$cost
```


# Explore Data

First scatter 

```{r}
plot(sales$date, sales$demand.ham, main="Ham")
plot(sales$date, sales$demand.turkey, main="Turkey")
plot(sales$date, sales$demand.veggie, main="Veggie")

plot(sales$date, sales$available.ham, main="Ham")
plot(sales$date, sales$available.turkey, main="Turkey")
plot(sales$date, sales$available.veggie, main="Veggie")
```

What does total demand look like?

```{r}
plot(sales$demand.total, ylim=c(0, 80))

# Does this depend on the day of the week?
m <- seq(from=1, to = 130, by = 5)
plot(sales$demand.total[m], main="monday")
plot(sales$demand.total[m+1], main="tuesday")
plot(sales$demand.total[m+2], main="wednesday")
plot(sales$demand.total[m+3], main="thursday")
plot(sales$demand.total[m+4], main="friday")

```

Let's look at some histograms

```{r}
hist(sales$demand.total[m], main="monday")
hist(sales$demand.total[m+1], main="tuesday")
hist(sales$demand.total[m+2], main="wednesday")
hist(sales$demand.total[m+3], main="thursday")
hist(sales$demand.total[m+4], main="friday")
```

That looks much better. Now let's look at sandwich by day of week

```{r}
#Ham

hist(sales$demand.ham[m], main="monday")
hist(sales$demand.ham[m+1], main="tuesday")
hist(sales$demand.ham[m+2], main="wednesday")
hist(sales$demand.ham[m+3], main="thursday")
hist(sales$demand.ham[m+4], main="friday")
```

```{r}
#Turkey

hist(sales$demand.turkey[m], main="monday")
hist(sales$demand.turkey[m+1], main="tuesday")
hist(sales$demand.turkey[m+2], main="wednesday")
hist(sales$demand.turkey[m+3], main="thursday")
hist(sales$demand.turkey[m+4], main="friday")
```

```{r}
# Veggie

hist(sales$demand.veggie[m], main="monday")
hist(sales$demand.veggie[m+1], main="tuesday")
hist(sales$demand.veggie[m+2], main="wednesday")
hist(sales$demand.veggie[m+3], main="thursday")
hist(sales$demand.veggie[m+4], main="friday")
```

```{r}
What about just demand per sandwich
hist(sales$demand.veggie)
hist(sales$demand.turkey)
hist(sales$demand.ham)
```

I think we can model demand as a poisson distribution where lambda will be a function of at least sandwich and weekday.

# Ham

Let's start looking at the Ham sandwich. I will begin by using a poisson distribution to model the demand for a Ham sandwich. The expected value = $\lambda$ = average Ham sandwich sold per day.

```{r}
lambda <- sum(sales$demand.ham) / nrow(sales)
ndays <- nrow(sales)

trials <- 10000
n <- 10000 / ndays
sim.sales <- rpois(trials, lambda)
barplot(table(sim.sales) / n)
hist(sales$demand.ham)

```

Now I need to model the profit:

profit = demand * profit - leftover * cost
inventory = 16

What happens?

```{r}
days <- 1000
stock <- 15
i = 1
results <- vector(mode="numeric")
for(stock in 10:20){
  demand <- rpois(days, lambda)
  profit <- sum(demand >= stock) * details$profit[details$type == "ham"] * stock
  loss <- sum((demand - 16)[demand - 16 < 0] * details$cost[details$type == "ham"])
  results[i] <- profit + loss
  i = i + 1
}


```



# Comments

Introduciton says 2 years but we only have data for 26 weeks. Why?

# TODO

* Next I'd like to look to see if availability affects this.


